\section{Scalability}\label{sec:scalability}
We conducted two scalability experiments with the \gls{lmac} protocol from \autoref{sec:lmacc}. The first
experiment is to show how simulations with the centralised Coordinator performs, with a fixed number of nodes,
when scaling the amount of CPUs available for the simulation. In the second experiment, we attempt to scale
the number of nodes, while also scaling the number of CPUs to match the size of the network.

\begin{table}[H]
    \begin{tabular}{|l|c|c|c|}
        \hline
        Cores & 16        & 32        & 64        \\ \hline
        Time  & 41:49 min & 34:52 min & 19:28 min \\ \hline
    \end{tabular}
    \caption{5-minute random walk topology, 60 nodes.}\label{table:lmac-experiment1}
\end{table}

In the first experiment, we simulate the \gls{lmac} protocol with 60 nodes in a generated random walk
topology, over a 5-minute real-time period. The results of this experiment can be seen in
\autoref{table:lmac-experiment1}. The experiment demonstrated that scaling the number of CPUs will
significantly improve the time required to run the simulation, but it also shows that the simulation does not
scale very well with a larger number of nodes, as even when running the simulation with 64 cores available,
simulating the experiment takes way longer than the 5-minute \gls{gps} log we simulate.
% 5-minute simulation, 60 nodes, random walk topology
% 16 cores : 41:49
% 32 cores : 34:52
% 64 cores : 19:28

\begin{table}[H]
    \begin{tabular}{|l|c|c|c|}
        \hline
        Nodes & 100       & 300         & 600   \\ \hline
        Cores & 128       & 320         & 576   \\ \hline
        Time  & 45:43 min & 6:51:53 hrs & \dots \\ \hline
    \end{tabular}
    \caption{5-minute random walk topology.}\label{table:lmac-experiment2}
\end{table}

The second experiment further shows the scalability issues of the Coordinator. In this experiment, we
attempted to simulate the \gls{lmac} protocol while scaling the number of nodes in the network, as well as the
number of CPUs available. The results of this experiment can be seen in \autoref{table:lmac-experiment2}. A
5-minute real-time simulation with just 100 nodes took over 45 minutes with 128 cores, and a 300 node
experiment, with 320 cores took almost seven hours. After this result, we decided not to continue the
experiment.

% 5-minute simulation, random walk topology
% 100 nodes : 45:43
% 300 nodes : 6:51:53
% 600 nodes :
% 1000 nodes : 
